### 2021-07-04 docs/manuscript/manuscript.tex
Rephrase text

Cells interact continuously with their environment and they must detect signals
from it. In many cases Gene Regulatory Networks (GRNs) interpret these signals
through gene interaction and produce a particular response. This information
processing function is of particular importance in developmental processes such
as those in which a group of cells differentiate in response to a signalling
molecule. Such molecules were referred to as morphogens by Turing in
1952~\cite{Turing1952} and posterior theoretical studies on patterning led to the
conceptualization of the French Flag Problem by Wolpert, who also stated
in this respect that a gradient of concentration of a morphogen could
trigger cell differentiation in a one-dimensional field of cells~\cite{Wolpert1969,
sharpe_2019}.
Although the Bicoid protein (bcd), the first example of a molecule that acted
as a morphogen, was only found in the 80's in the developing embryos of
\textit{Drosophila melanogaster}~\cite{driever_gradient_1988,
driever_bicoid_1988}, now it is known that there
are many other examples such as the Decapentaplegic (Dpp) protein in
\textit{Drosophila} wing imaginal discs~\cite{affolter_decapentaplegic_2007},
as well as Sonic Hedgehog~\cite{dessaud_dynamic_2010,cohen_theoretical_2014}
and Wnt~\cite{raspopovic_digit_2014}. However, information processing by GRNs
has proven to be a complex process, and a major goal of developmental biology
is to understand mechanistically how positional information given by morphogens
is translated in spatial differentiation.\\

@@

The optimal pattern of gene expression defined by us, consisted in cells at the
sides of the field ($n<11$ and $n>20$) displaying expression levels lower than
10\% of the maximal level observed along the field of cells for the output gene,
and cells at the middle of the field ($n ∈ [11,20]$) displaying expression
levels greater than 90\% of the maximal level observed along the entire field
for the output gene (Fig~\ref{fig:model}C).

@@

The degree distribution of this network seems to follow a Poisson distribution
(Pearson’s $\chi^2 = 12.8902$, $\text{p-value} = 0.115684$) characteristic of a
Erdös-Rényi network with many nodes and edges~\cite{Erdos1959}. This kind of
network presents a great
number of nodes with low degree and a few nodes with high degree
(Fig~\ref{fig:deg-dist}A). Nodes corresponding to topologies number one and
three showed the highest degree ($deg(1) = deg (3) = 14$). Curiously, nodes with
higher degree resulted to be also those network topologies with higher abundance
(Fig~\ref{fig:deg-dist}B). This relationship between abundance and node degree
was corroborated by the Spearman's rank correlation coefficient
($r_S = 0.353752$, $p = 6.2933\times10^{-23} $).

@@

We wanted to know if connecting topologies separated by two mutational steps
would produce a connected graph, for that reason we generated another neutral
network joining neighbors at two mutational steps and we found that only
topology number 420 remained unconnected from the rest of topologies in the
graph (Fig~\ref{fig:2neut-net}).

@@

Besides topological robustness, we wanted to know if network topologies were
robust to perturbations in the values of the parameters. After performing the
perturbations to a GRN from each topology, we plotted the abundance of the
topology vs. its robustness. We did not find a significant correlation between
these two variables ($r_S = −0.0389724$, $\text{p-value} = 0.298475$)
(Fig~\ref{fig:ab-rob}A). However, when we grouped those topologies containing
the same subgraph and calculated their mean robustness, we found that between
the most robust subgraphs is the ``Bistable'' motif reported by Cotterell \&
Sharpe~\cite{Cotterell2010}, the I3-FFL and the ``Overlapping domains'' motif
(which is a variation of the I3-FFL) also reported by Cotterell \& Sharpe
(see \nameref{S2_Fig} and \nameref{S1_Table}).

>>>>

Cells interact continuously with their environment, which requires very specific
regulatory strategies to avoid potential detrimental responses. Arguably, most
cellular functions arise from the dynamic activity of Gene Regulatory Networks
(GRNs), which play a central role in the interpretation of external and internal
signals. This information
processing function is of particular importance in developmental processes such
as those in which a group of cells differentiate in response to a signalling
molecule. Such molecules were referred to as morphogens by Turing in
1952~\cite{Turing1952} and posterior theoretical studies on patterning led to the
conceptualization of the French Flag Problem by Wolpert, who also stated
in this respect that a gradient of concentration of a morphogen could
trigger cell differentiation in a one-dimensional field of cells~\cite{Wolpert1969,
sharpe_2019}.
Although the Bicoid protein (bcd), the first example of a molecule that acted
as a morphogen, was only found in the 80's in the developing embryos of
\textit{Drosophila melanogaster}~\cite{driever_gradient_1988,
driever_bicoid_1988}, now it is known that there
are many other examples such as the Decapentaplegic (Dpp) protein in
\textit{Drosophila} wing imaginal discs~\cite{affolter_decapentaplegic_2007},
as well as Sonic Hedgehog~\cite{dessaud_dynamic_2010,cohen_theoretical_2014}
and Wnt~\cite{raspopovic_digit_2014}. However, information processing by GRNs
has proven to be a complex process, and a major goal of developmental biology
is to understand mechanistically how positional information conveyed by morphogens
is translated into spatial differentiation.\\

@@

The optimal pattern of gene expression defined in this study consists in cells at
the border of the field ($n<11$ and $n>20$) displaying expression levels lower than
10\% of the maximal level observed along the field of cells for the output gene,
and cells at the middle of the field ($n ∈ [11,20]$) displaying expression
levels greater than 90\% of the maximal level observed along the entire field
for the output gene (Fig~\ref{fig:model}C).

@@

The degree distribution of this network seems to follow a Poisson distribution
(Pearson’s $\chi^2 = 12.8902$, $\text{p-value} = 0.115684$) characteristic of a
Erdös-Rényi network with many nodes and edges~\cite{Erdos1959}. This kind of
network presents a great
number of nodes with low degree and a few nodes with high degree
(Fig~\ref{fig:deg-dist}A). Nodes corresponding to topologies number one and
three showed the highest degree ($deg(1) = deg (3) = 14$). Curiously, we found
that nodes with higher degree turned out to be also those network topologies
with higher abundance
(Fig~\ref{fig:deg-dist}B). This relationship between abundance and node degree
was corroborated by the Spearman's rank correlation coefficient
($r_S = 0.353752$, $p = 6.2933\times10^{-23} $).

@@

Furthermore, we examined the extent to which the metagraph of topologies
remained highly connected when joining neighbors at two mutational steps.
Interestingly, we found that only topology number 420 remained isolated from
the rest of topologies in the metagraph (Fig~\ref{fig:2neut-net}).

@@

Next to topological robustness, we investigated the robustness of the sampled
network topologies to parameter changes (see Methods). Fig~\ref{fig:ab-rob}A
illustrates the association between the abundance of a topology class versus its
robustness. We did not find a significant correlation between
these two variables ($r_S = −0.0389724$, $\text{p-value} = 0.298475$).
However, when we grouped those topologies containing
the same subgraph and calculated their mean robustness, we found that between
the most robust subgraphs is the ``Bistable'' motif reported by Cotterell \&
Sharpe~\cite{Cotterell2010}, the I3-FFL and the ``Overlapping domains'' motif
(which is a variation of the I3-FFL) also reported by Cotterell \& Sharpe
(see \nameref{S2_Fig} and \nameref{S1_Table}).
